### 2021-09-16 ReadingGroupPres.tex
Updates from Overleaf

\documentclass[10pt]{beamer}

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage[style=authoryear]{biblatex}
\bibliography{demo.bib}
\usepackage{pgfplots}

\usepgfplotslibrary{dateplot}

\usepackage{xspace}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}

\title{Hospital Information Technology and Physician Response}

\subtitle{Hanna Glenn}
 \date{\today}
% \date{}
%\author{Presented by: Hanna Glenn}
% \institute{Center for modern beamer themes}
% \titlegraphic{\hfill\includegraphics[height=1.5cm]{logo.pdf}}

\begin{document}

\maketitle

\setbeamercolor{background canvas}{bg=white}

\begin{frame}{Table of Contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents%[hideallsubsections]
\end{frame}

\section[Motivation]{Motivation}

\begin{frame}{The Beginning of HIT}
\centering
    \includegraphics[scale=.6]{graphics/News Clip4.PNG}
\end{frame}

\begin{frame}[fragile]{HIT has had Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item A 2005 estimate finds possible cost reduction of hundreds of billions of dollars (\cite{hillestad2005})
\end{itemize}
\end{alertblock}

\begin{alertblock}{Quality Improvement}
\begin{itemize}
    \item Improved efficiency, patient safety improvements, physicians have decision support that could prevent unnecessary complications, etc.
    \item Significant policy push for EHR implementation with these goal in mind: HITECH Act, 2008 provided financial incentive for hospitals to implement EHRs \nocite{hitech}
\end{itemize}
\end{alertblock}

\textcolor{blue}{The percentage of hospitals with basic EHR capability rose from 9$\%$ in 2008 to 84$\%$ in 2015.} (\cite{stats})

\end{frame}

\begin{frame}[fragile]{What does this mean for physicians?}
\begin{itemize}
    \item Day-to-day, this means physicians spend more time entering information into a computer 
    \item Loss of autonomy as EHRs have progressed
    \item Physicians have experienced significant burn-out from EHRs
\end{itemize}

\end{frame}

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.4]{graphics/News Clip1.PNG}
\end{center}
\end{frame}

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.5]{graphics/News Clip2.PNG}
\end{center}
\end{frame}

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.3]{graphics/News Clip3.PNG}
\end{center}
\end{frame}

\begin{frame}{This Paper}

Does EHR implementation in hospitals affect labor market decisions of physicians, specifically where, and how much they work?

\end{frame}

\section{Contribution}

\begin{frame}{Branches of Literature}
\centering
\includegraphics[scale=.45]{graphics/Contribution_litgraphic.PNG}

\end{frame}

\begin{frame}{Physician Labor Literature}
\centering
    \includegraphics[scale=.45]{graphics/labor_litgraphic.PNG}
\end{frame}

\begin{frame}{EHR Literature}
    \centering
    \includegraphics[scale=.45]{graphics/EHR_litgraphic.PNG}
\end{frame}

\section{Model and Data}

\begin{frame}{Physician's Incentives}
\begin{itemize}
    \item Physician decides to work, possibly at a handful of hospitals
    \item Some proportion of these implement a new technology (an EHR)
    \item Three ways to think about the effect:
    \begin{enumerate}
        \item The EHR could impose a burden (cost) on the physician, leading them to shift work to other hospitals or lessen their own workload
        \item The EHR could increase the physician's attachment to the implementing hospital (through the learning of a complicated new skill), leading them to lessen work at other hospitals and increase work at EHR hospital
        \item Other factors impact physicians more than the hospital's technology, so they do not change their behavior when EHRs are implemented
    \end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}{Relevant Data}
    \begin{equation*}
    p_{iht}=\textbf{p}(EHR_{ht},EHR_{-h,t},X_i,H_j,...)
\end{equation*}
\end{frame}

\begin{frame}[noframenumbering]{Relevant Data}
\begin{equation*}
    \textcolor{blue}{p_{iht}}=\textbf{p}(EHR_{ht},EHR_{-h,t},X_i,H_j,...)
\end{equation*}

$\textcolor{blue}{p_{iht}}:$
\begin{itemize}
    \item measures of physician's labor
    \item number of shared patients with a hospital: CMS PSPD (2009-2015)
    \item total services: Medicare Provider Utilization and Payment Data (starts in 2012)
\end{itemize}
\end{frame}

\begin{frame}[noframenumbering]{Relevant Data}
\begin{equation*}
    p_{iht}=\textbf{p}(\textcolor{red}{EHR_{ht},EHR_{-h,t}},X_i,H_j,...)
\end{equation*}

$\textcolor{red}{EHR}:$
\begin{itemize}
    \item Main variable (binary for whether a hospital uses an EHR) comes from AHA survey
    \item AHA IT Supplement also has information on EHRs that can help with documentation or decision making
    \item Meaningful use data gives information on which hospitals received a subsidy for achieving "meaningful use"
\end{itemize}
\end{frame}

\begin{frame}[noframenumbering]{Relevant Data}
\begin{equation*}
    p_{iht}=\textbf{p}(EHR_{ht},EHR_{-h,t},\textcolor{purple}{X_i},H_j,...)
\end{equation*}

$\textcolor{purple}{X_i}:$
\begin{itemize}
    \item Physician level characteristics
    \item Medical school graduation date, gender: Physician Compare
\end{itemize}
\end{frame}

\begin{frame}[noframenumbering]{Relevant Data}
\begin{equation*}
    p_{iht}=\textbf{p}(EHR_{ht},EHR_{-h,t},X_i,\textcolor{orange}{H_{jt}},...)
\end{equation*}

$\textcolor{orange}{H_{jt}}:$
\begin{itemize}
    \item Hospital Level Characteristics
    \item Number of beds: AHA Survey
    \item Number of days out of the year operating: AHA Survey
\end{itemize}
\end{frame}

\begin{frame}{Summary Statistics}
\centering
\includegraphics[scale=.45]{Objects/sumstats_physician_table.pdf}\\
\vspace{3mm}
\includegraphics[scale=.45]{Objects/sumstats_pair_table.pdf}\\
\vspace{3mm}
\includegraphics[scale=.45]{Objects/sumstats_hospital_table.pdf}
\end{frame}

\begin{frame}{Variation in EHR Adoption}
\centering
\includegraphics[scale=.55]{Objects/TYP_plot_hospEHR_year.pdf}
\end{frame}

\begin{frame}{Variation in Working Variables}
    \centering
    \includegraphics[scale=.5]{Objects/mainhosp_share_treatvscontrol.pdf}
\end{frame}

\begin{frame}{Variation in Meaningful Use}
    \centering
    \includegraphics[scale=.5]{Objects/TYP_plot_hospmeanuse_year.pdf}
\end{frame}

\begin{frame}{Variation in Working Variables}
\centering
\includegraphics[scale=.5]{Objects/hhi_treatvscontrol.pdf}
\end{frame}

\begin{frame}{Variation in Working Variables}
    \centering
    \includegraphics[scale=.4]{Objects/rel_mainhosp_share_treatvscontrol.pdf}
\end{frame}

\section{Plans}

\begin{frame}{Estimation}
First, DD
\begin{itemize}
    \item Parallel trends assumption
    \item EHR implementation exogenous to physician?
    \item Implement all the current DD techniques
\end{itemize}

Later, IV
\begin{itemize}
    \item State level HIPPA privacy laws
    \item Bandwidth capacity of a given area
\end{itemize}
\end{frame}

\begin{frame}{Other Plans}
\begin{itemize}
    \item Dig deeper into meaningful use measures
    \item Try to get location measurements
\end{itemize}

\end{frame}

\end{document}