### 2021-10-28 Presentations/ThirdYearPaperpres.tex
Updates from Overleaf

\documentclass[10pt]{beamer}

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{pgfpages}
\usepackage{booktabs}
\usepackage{xcolor}

\usepackage{pgfplots}

\usepgfplotslibrary{dateplot}

\setbeameroption{slides only}
%\setbeameroption{show notes on second screen=right}
\usepackage{xspace}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}

\title{Labor Markets and Technological Change: Evidence from Electronic Health Records}

\subtitle{Hanna Glenn}
 \date{\today}

\setbeamertemplate{itemize item}{\scalebox{.6}{$\blacktriangleright$ }}     
\setbeamertemplate{itemize subitem}{\scalebox{.8}{$\centerdot$ }}

\begin{document}

\maketitle

\setbeamercolor{background canvas}{bg=white}

\begin{frame}{Table of Contents}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents%[hideallsubsections]

  \note[item]{Roadmap of the presentation}
\end{frame}

\section[Motivation]{Motivation}

\begin{frame}{Technology and Displacement}

\end{frame}

\begin{frame}[fragile]{What is an Electronic Health Record?}
\begin{itemize}
    \item Technology purchased and implemented by hospitals
\end{itemize}

\begin{itemize}
    \item Used for:
    \vspace{3mm}
    \begin{itemize}
        \item Electronic patient records
        \vspace{3mm}
        \item Decision making assistance
    \end{itemize}
\end{itemize}

\note[item]{Get everyone on the same page about how to think about EHRs}
\note[item]{A product in the form of technology that is sold to hospitals (and others, but we focus on hospital setting)}
\note[item]{The most standard element of an EHR is that it converts paper records to electronic records}
\note[item]{New elements of EHRs: interropability, patient interface, decision-making assistance}

\end{frame}

\begin{frame}[fragile]{HIT: Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item Possible cost reduction of hundreds of billions of dollars \\ \scriptsize (Hillestad et al 2005)
\end{itemize}
\end{alertblock}

\begin{alertblock}{Quality Improvement}
\begin{itemize}
    \item Improved efficiency, patient safety improvements, physicians have decision support that could prevent unnecessary complications, etc.
    \item Significant policy push for EHR implementation: HITECH Act, 2008 provided financial incentive for hospitals to implement EHRs 
\end{itemize}
\end{alertblock}

\vspace{4mm}

\textcolor{blue}{The percentage of hospitals with basic EHR capability rose from 9$\%$ in 2008 to 84$\%$ in 2015.} \scriptsize (Health IT Dashboard)

\note[item]{Why do we care about EHRs or health information technology in the first place?}
\note[item]{Verbally explain the bridge to physicians here (the slide I took out)}
\end{frame}

\begin{frame}[noframenumbering]{Physician Response to EHRs}
\begin{center}
    \includegraphics[scale=.5]{graphics/News Clip2.PNG}
\end{center}
\end{frame}

\begin{frame}[noframenumbering]{Physician Response to EHRs}
\begin{center}
    \includegraphics[scale=.35]{graphics/News Clip3.PNG}
\end{center}
\end{frame}

\begin{frame}{This Paper}

Did EHR implementation in hospitals affect physician labor market decisions?
\begin{itemize}
    \item Intensive Margin / Productivity
    \item Extensive Margin
    \item Hospital $\rightarrow$ Office
\end{itemize}
\end{frame}

\section{Contribution}

\begin{frame}{Literature}
\footnotesize 
\begin{columns}
    \column{0.34\textwidth}
        \centering
        \underline{ EHR $\rightarrow$ Healthcare }
        \begin{itemize}
            \item Small improvements for severe health conditions \\ \vspace{1mm}
            \tiny(Agha 2014; McCullough et al 2016)

            \footnotesize

            \item Higher ``levels" of EHR use $\rightarrow$ decline in quality of care\\ \vspace{1mm}
            \tiny (Appari et al 2013)

            \footnotesize

            \item No decrease in cost \\ \vspace{1mm}
            \tiny (Agha 2014; Dranove et al 2019)

            \footnotesize

            \item Productivity and opinion about EHRs varies drastically \\ \vspace{1mm}
            \tiny (Hitt $\&$ Tambe 2016, Butler $\&$ Johnson 2016, Meyerhoefer et al 2016)

        \end{itemize}

    \column{0.33\textwidth}
    \color{gray}
        \centering
        \underline{ Technology $\rightarrow$ Labor }
        \begin{itemize}
        \color{gray}
            \item Erosion effect vs. wage effect \\ \vspace{1mm}
            \tiny (Zeira $\&$ Joseph 2011)

            \footnotesize

            \item Cause retirement? \\ \vspace{1mm}
            \tiny (Schleife 2006, Cavapozzi et al 2013, Roger et al 2006, Friedburg 2003)

            \footnotesize

            \item Older, well educated members of labor force exited at higher rates due to computerization \\ \vspace{1mm}
            \tiny (Willis and Hudomiet 2020)
        \end{itemize}

    \column{0.33\textwidth}
        \color{gray}
        \centering
        Physician Labor \underline{ Determinants }
        \begin{itemize}
        \color{gray}
            \item Physicians not very responsive to wage changes in terms of hours worked \\ \vspace{1mm}
            \tiny (Baltagi et al 2005, Sloan 2018)

            \footnotesize

            \item Peronal life factors matter more for retirement than work conditions \\ \vspace{1mm}
            \tiny (Bahrami 2002)

            \footnotesize

            \item Physician burnout: early retirement and less hours worked \\ \vspace{1mm} 
            \tiny (Dewa et al 2014)
        \end{itemize}
\end{columns}

\note[item]{\footnotesize Agha 2014 (JHE) uses claims data and difference in difference methodology. Find 1.3 percent increase in billed costs and no improvements in health coutcomes, data is 1998-2005.}

\note[item]{McCullough, Parente, Town 2016, RAND. claims data 2002-2007. Find no reduction in moretality for the median patient, only those who require complex care with coordination.}

\note[item]{Appari, Johnson, Anthony 2013. Health services review. fixed effects linear panel model. quality measure comes from hospital compare, not at the patient level.}

\note[item]{Dranove, Forman, Goldfarb, Greenstein (2019) American Economic Journal: Economic Policy. data from 1996 to 2009. Specific findings: hospitals have a rise in cost after 3 years, some hospitals in certain locations have a decrease in costs after 5 years but others still have a higher cost after 6 years.}

\end{frame}

\begin{frame}[noframenumbering]{Literature}
\footnotesize 
\begin{columns}
    \column{0.34\textwidth}
    \color{gray}
        \centering
        \underline{ EHR $\rightarrow$ Healthcare }
        \begin{itemize}
        \color{gray}
            \item Small improvements for severe health conditions \\ \vspace{1mm}
            \tiny(Agha 2014, JHE; McCullough et al 2016, RAND; Meyerhoefer et al, ILRR)

            \footnotesize

            \item Higher ``levels" of EHR use $\rightarrow$ decline in quality of care\\ \vspace{1mm}
            \tiny (Appari et al 2013)

            \footnotesize

            \item No decrease in cost \\ \vspace{1mm}
            \tiny (Agha 2014, JHE; Dranove et al 2019; AEJ)

            \footnotesize

            \item Productivity and opinion about EHRs varies drastically \\ \vspace{1mm}
            \tiny (Hitt $\&$ Tambe 2016, Butler $\&$ Johnson 2016, Meyerhoefer et al 2016)

        \end{itemize}

    \column{0.33\textwidth}
        \centering
        \underline{ Technology $\rightarrow$ Labor }
        \begin{itemize}
            \item Erosion effect vs. wage effect \\ \vspace{1mm}
            \tiny (Zeira $\&$ Joseph 2011)

            \footnotesize

            \item Cause retirement? \\ \vspace{1mm}
            \tiny (Schleife 2006, Cavapozzi et al 2013, Roger et al 2006, Friedburg 2003)

            \footnotesize

            \item Older, well educated members of labor force exited at higher rates due to computerization \\ \vspace{1mm}
            \tiny (Willis and Hudomiet 2020)
        \end{itemize}

    \column{0.33\textwidth}
    \color{gray}
        \centering
        Physician Labor \underline{ Determinants }
        \begin{itemize}
        \color{gray}
            \item Physicians not very responsive to wage changes in terms of hours worked \\ \vspace{1mm}
            \tiny (Baltagi et al 2005, Sloan 2018)

            \footnotesize

            \item Peronal life factors matter more for retirement than work conditions \\ \vspace{1mm}
            \tiny (Bahrami 2002)

            \footnotesize

            \item Physician burnout: early retirement and less hours worked \\ \vspace{1mm} 
            \tiny (Dewa et al 2014)
        \end{itemize}
\end{columns}

\note[item]{Zeira and Johnson 2011. }
\end{frame}

\begin{frame}[noframenumbering]{Literature}
\footnotesize 
\begin{columns}
    \column{0.34\textwidth}
    \color{gray}
        \centering
        \underline{ EHR $\rightarrow$ Healthcare }
        \begin{itemize}
        \color{gray}
            \item Small improvements for severe health conditions \\ \vspace{1mm}
            \tiny(Agha 2014, JHE; McCullough et al 2016, RAND; Meyerhoefer et al, ILRR)

            \footnotesize

            \item Higher ``levels" of EHR use $\rightarrow$ decline in quality of care\\ \vspace{1mm}
            \tiny (Appari et al 2013)

            \footnotesize

            \item No decrease in cost \\ \vspace{1mm}
            \tiny (Agha 2014, JHE; Dranove et al 2019; AEJ)

            \footnotesize

            \item Productivity and opinion about EHRs varies drastically \\ \vspace{1mm}
            \tiny (Hitt $\&$ Tambe 2016, Butler $\&$ Johnson 2016, Meyerhoefer et al 2016)

        \end{itemize}

    \column{0.33\textwidth}
    \color{gray}
        \centering
        \underline{ Technology $\rightarrow$ Labor }
        \begin{itemize}
        \color{gray}
            \item Erosion effect vs. wage effect \\ \vspace{1mm}
            \tiny (Zeira $\&$ Joseph 2011)

            \footnotesize

            \item Cause retirement? \\ \vspace{1mm}
            \tiny (Schleife 2006, Cavapozzi et al 2013, Roger et al 2006, Friedburg 2003)

            \footnotesize

            \item Older, well educated members of labor force exited at higher rates due to computerization \\ \vspace{1mm}
            \tiny (Willis and Hudomiet 2020)
        \end{itemize}

    \column{0.33\textwidth}
        \centering
        Physician Labor \underline{ Determinants }
        \begin{itemize}
            \item Physicians not very responsive to wage changes in terms of hours worked \\ \vspace{1mm}
            \tiny (Baltagi et al 2005, Sloan 2018)

            \footnotesize

            \item Peronal life factors matter more for retirement than work conditions \\ \vspace{1mm}
            \tiny (Bahrami 2002)

            \footnotesize

            \item Physician burnout: early retirement and less hours worked \\ \vspace{1mm} 
            \tiny (Dewa et al 2014)
        \end{itemize}
\end{columns}
\end{frame}

\section{Data}

\begin{frame}{About the Data}
    Main data: \underline{CMS Shared Patient Data} (2009-2015)
    \begin{itemize}
        \item Lists 2 NPIs and the number of billings those two entities shared
    \end{itemize}
    \vspace{3mm}

    How I use it:
    \begin{itemize}
        \item Physician-hospital pairs (PCPs, internists, hospitalists)
        \item Billings that occur on same day
        \item Close relationship between hospital and physician
        \item Aggregate to the physician level
    \end{itemize}
\end{frame}

\begin{frame}{Measures of Physician Labor Market Decisions}

     Patients visited in hospitals

     \vspace{4mm}

     Indicator for switching work setting:
    \begin{itemize}
        \item 0 patients visited but maintaining billable activity
    \end{itemize}

    \vspace{4mm}

    \textcolor{gray}
     {Indicator for retirement/ leaving the labor force}

\end{frame}

\begin{frame}{Measure of EHR Use}

Physician-level treatment variable that captures exposure to \textit{any} EHR
\vspace{3mm}
\begin{itemize}
    \item Individual hospital EHR use comes from AHA survey
    \vspace{3mm}
    \item Treatment turns on the first time any of the physicians' hospitals implements an EHR
\end{itemize}
\end{frame}

\begin{frame}{Other Characteristics}

Medical school graduation date: \underline{Physician Compare}

\end{frame}

\begin{frame}{Data}
\centering
    \includegraphics[scale=.8]{Objects/sumstats.pdf}

\end{frame}

\begin{frame}{Physician EHR Use}
    \centering
    \includegraphics[scale=.5]{Objects/sum_stats_year.pdf}
\end{frame}

\section{Analysis}

\begin{frame}{Event Study}
\begin{equation*}
    y_{it}=\alpha_i+\delta_t+q_{it}'\lambda+\sum_{k=-5}^5 \beta_kz_{i,t-k} + \varepsilon_{it}
\end{equation*}

\end{frame}

\section{Results}

\begin{frame}{Dep. Variable: Patients Billed with Hospitals}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_fullsample.pdf}
            \caption{\small All Physicians\\}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_oldsample.pdf}
            \caption{\small Phys. $>$ 35 yrs Experience}
        \end{minipage}
    \end{figure}
\end{frame}

\begin{frame}{Dep. Variable: Patients Billed with Hospitals}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_hosp_fullsample.pdf}
            \caption{\small All Physicians \\(Only 1 Hospital)}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_hosp_oldsample.pdf}
            \caption{\small Phys. $>$ 35yrs\\ (Only 1 Hospital)}
        \end{minipage}
    \end{figure}
\end{frame}

\begin{frame}{Dep. Variable: Probability of Billing, but not With Hospitals}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_fullsample_ind.pdf}
            \caption{\small All Physicians}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_oldsample_ind.pdf}
            \caption{\small Phys. $>$ 35yrs Experience}
        \end{minipage}
    \end{figure}
\end{frame}

\begin{frame}{Dep. Variable: Probability of Billing, but not With Hospitals}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_hosp_fullsample_ind.pdf}
            \caption{\small All Physicians \\(Only 1 Hospital)}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_hosp_oldsample_ind.pdf}
            \caption{\small Phys. $>$ 35yrs Experience\\ (Only 1 Hospital)}
        \end{minipage}
    \end{figure}
\end{frame}

\begin{frame}{Dep. Variable: Appears in Office-Based Data}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_fullsample_office.pdf}
            \caption{\small All Physicians}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_oldsample_office.pdf}
            \caption{\small Phys.$>$35yrs Experience}
        \end{minipage}
    \end{figure}
\end{frame}

\begin{frame}{Dep. Variable: Appears in Office-Based Data}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_hosp_fullsample_office.pdf}
            \caption{\small All Physicians \\(Only 1 Hospital)}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_hosp_oldsample_office.pdf}
            \caption{\small Phys.$>$35yrs Experience\\ (Only 1 Hospital)}
        \end{minipage}
    \end{figure}
\end{frame}

\section{Continuing Work}

\begin{frame}{Plans}
    \begin{itemize}
        \item Analysis on retirement decision of physicians: requires a better understanding of why physicians would drop out of this Medicare data
        \item Investigate other estimation strategies
    \end{itemize}
\end{frame}

\end{document}