### 2021-11-01 Presentations/ThirdYearPaperpres.tex

\documentclass[10pt]{beamer}

\usetheme[progressbar=frametitle]{metropolis}
\usepackage{appendixnumberbeamer}
\usepackage{pgfpages}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{mathtools}

\usepackage{pgfplots}

\usepgfplotslibrary{dateplot}

\setbeameroption{slides only}
%\setbeameroption{show only notes}
%\setbeameroption{show notes on second screen=right}
\usepackage{xspace}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}

\title{Labor Markets and Technological Change: Evidence from Electronic Health Records}

\subtitle{Hanna Glenn}
 \date{\today}

\setbeamertemplate{caption}{\raggedright\insertcaption\par}
\setbeamertemplate{itemize item}{\scalebox{.6}{$\blacktriangleright$ }}     
\setbeamertemplate{itemize subitem}{\scalebox{.8}{$\centerdot$ }}

\begin{document}

\maketitle

\setbeamercolor{background canvas}{bg=white}

\section[Motivation]{Motivation}

\begin{frame}{The Broad Question}
\begin{itemize}
    \item Longstanding topic in economics: relationship between technology and employment
    \vspace{2mm}
    \begin{itemize}
        \item Displacement?
        \item Productivity?
        \item Wages?
    \end{itemize}
    \vspace{3mm}
    \item Investigated in many settings
    \vspace{3mm}
    \item This paper: specific healthcare setting
\end{itemize}

\end{frame}

\begin{frame}[fragile]{What is an Electronic Health Record?}
``An electronic health record is the systematized collection of patient and population electronically stored health information in a digital format"

\vspace{3mm}

What it can do:
\begin{itemize}
    \item Paper $\rightarrow$ computerization
    \item Communication
    \item Decision making assistance
\end{itemize}

\end{frame}

\begin{frame}[fragile]{HIT: Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item Possible cost reduction of hundreds of billions of dollars \\ \scriptsize (Hillestad et al 2005)
\end{itemize}
\end{alertblock}

\begin{alertblock}{Quality Improvement}
\begin{itemize}
    \item Improved efficiency, patient safety improvements, physicians have decision support that could prevent unnecessary complications, etc.
    \item Significant policy push for EHR implementation: HITECH Act, 2008 provided financial incentive for hospitals to implement EHRs 
\end{itemize}
\end{alertblock}

\vspace{4mm}

\textcolor{blue}{The percentage of hospitals with basic EHR capability rose from 9$\%$ in 2008 to 84$\%$ in 2015.} \scriptsize (Health IT Dashboard)

\end{frame}

\begin{frame}{Physician Response to EHRs}
\begin{center}
    \includegraphics[scale=.3]{graphics/News Clip3.PNG}

    \vspace{3mm}

    \includegraphics[scale=.3]{graphics/News Clip2.PNG}
\end{center}

\end{frame}

\begin{frame}{This Paper}

Did EHR implementation in hospitals affect physician labor market decisions?
\begin{itemize}
    \item Intensive Margin / Productivity
    \item Extensive Margin
    \item Hospital $\rightarrow$ Office
\end{itemize}

\end{frame}

\section{Contribution}

\begin{frame}{Literature}
\small
\begin{columns}
\setlength{\tabcolsep}{-5pt}
    \column{0.4\textwidth}
        \centering
        \underline{ Technology $\rightarrow$ Labor }
        \vspace{-1mm}
        \begin{itemize}
            \item Erosion effect vs. wage effect \\ \vspace{1mm}
            \tiny (Zeira $\&$ Joseph 2011, EJ)

            \footnotesize

            \item Cause retirement? \\ \vspace{1mm}
            \tiny (Schleife 2006, Labour; Cavapozzi et al 2013, AE; Roger et al 2006, EJ; Friedburg 2003, ILRR)

            \footnotesize

            \item Older, well educated members of labor force exited at higher rates due to computerization \\ \vspace{1mm}
            \tiny (Willis and Hudomiet 2021, Labour)
        \end{itemize}

        \column{0.4\textwidth}
        \centering
        \color{gray}
        \underline{ EHR $\rightarrow$ Healthcare }
        \vspace{-1mm}
        \begin{itemize}
        \color{gray}
            \item Small improvements for severe health conditions \\ \vspace{1mm}
            \tiny(Agha 2014, JHE; McCullough et al 2016, RAND)

            \footnotesize

            \item Higher ``levels" of EHR use $\rightarrow$ decline in quality of care\\ \vspace{1mm}
            \tiny (Appari et al 2013, HSR)

            \footnotesize

            \item No decrease in cost \\ \vspace{1mm}
            \tiny (Agha 2014, JHE; Dranove et al 2019, AEJ)

            \footnotesize

            \item Productivity and opinion about EHRs varies drastically \\ \vspace{1mm}
            \tiny (Hitt $\&$ Tambe 2016, ILRR; Butler $\&$ Johnson 2016, IJHEM; Meyerhoefer et al 2016, ILRR)

        \end{itemize}

\end{columns}

\end{frame}

\section{Data}

\begin{frame}{Data}
    \underline{Main: CMS Shared Patient Data (2009-2015)}
    \begin{itemize}
        \item Hospital-physician pairs: bill for same patient under Medicare
        \item Physician (PCP, hospitalist, internist) works in the hospital
    \end{itemize}

    \vspace{3mm}

    \underline{EHR Use:}
    \begin{itemize}
        \item AHA Survey
    \end{itemize}

    \vspace{3mm}

    \underline{Other:}
    \begin{itemize}
        \item Physician Compare
    \end{itemize}

    \vspace{3mm}

    Aggregated to physician level
\end{frame}

\begin{frame}{Variables}
Labor Market Outcomes:
\vspace{-2mm}
\begin{itemize}
    \item Patients visited in hospitals
    \item \color{gray} Maintaining billable activity only outside of hospitals
    \item \color{gray} Level of billable activity outside of hospitals
    \item \color{gray} Retirement/leaving the labor force
\end{itemize}

EHR Use:
\vspace{-2mm}
\begin{itemize}
    \item Indicator for any exposure to EHR
    \item \color{gray} Fraction of hospitals exposed to EHR
\end{itemize}

Other:
\vspace{-2mm}
\begin{itemize}
    \item Years of experience
    \item Hospital characteristics
\end{itemize}

\end{frame}

\begin{frame}{Data}
    \centering
    \includegraphics[scale=.5]{Objects/sum_stats_year.pdf}
\end{frame}

\begin{frame}{Data}
\centering
    \includegraphics[scale=.5]{Objects/cont_treatment_graph.pdf}

\end{frame}

\section{Analysis}

\begin{frame}{Event Study}
\begin{equation*}
    y_{it}=\alpha_i+\delta_t+q_{it}'\lambda+\sum_{k=-5}^5 \beta_kz_{i,t-k} + \varepsilon_{it}
\end{equation*}

\vspace{4mm}

\begin{align*}
    y_{it} &: \text{Outcomes} \hspace{30mm}\\
    \alpha_i, \delta_t &: \text{Fixed Effects}\hspace{30mm}\\
    q_{it} &: \text{Physician/hospital characteristics}\hspace{30mm}\\
    z_{i,t-k} &: \text{Exposure to EHR} \hspace{30mm}
\end{align*}

\end{frame}

\section{Results}

\begin{frame}{Dep. Variable: Patients Billed with Hospitals}
\begin{figure}[ht]
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_fullsample.pdf}
            \caption{\small All Physicians\\}
        \end{minipage}
        \hspace{0.2cm}
        \begin{minipage}[b]{0.47\linewidth}
            \centering
            \includegraphics[width=\textwidth]{Objects/xtevent_oldsample.pdf}
            \caption{\small Phys. $>$ 35 yrs Experience}
        \end{minipage}
    \end{figure}
\end{frame}

\section{Continuing Work}

\begin{frame}{Plans}
    \begin{itemize}
        \item Analysis on retirement decision of physicians: requires a better understanding of why physicians would drop out of this Medicare data
        \item Investigate other estimation strategies
    \end{itemize}
\end{frame}

\begin{frame}{}
\centering
    Thank you!
\end{frame}

\end{document}