### 2022-06-03 Table Code/split stats.tex

\begin{table}[h]

\caption{Means by Age Sample}
\centering
\begin{tabular}[t]{lccc}
\toprule
Variable & Age $>$ 60 & Age $<=$ 60 & Any Who Retire\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Outcomes}}\\
\hspace{1em}Ever Retire & 0.04 & 0.03 & 1.00\\
\hspace{1em}Fraction Patients in Office & 0.15 & 0.09 & 0.06\\
\hspace{1em}Work in an Office & 0.48 & 0.28 & 0.18\\
\hspace{1em}Change Zip Codes & 0.10 & 0.12 & 0.09\\
\hspace{1em}Number of Patients & 341.34 & 328.12 & 76.16\\
\hspace{1em}Claim Count & 1762.41 & 1330.00 & 290.68\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Treatment}}\\
\hspace{1em}Exposure to an EHR & 0.76 & 0.75 & 0.75\\
\hspace{1em}Fraction of Hospitals with EHR & 0.59 & 0.64 & 0.65\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Characteristics}}\\
\hspace{1em}Age & 60.31 & 42.75 & 45.96\\
\hspace{1em}Female & 0.18 & 0.36 & 0.40\\
\hspace{1em}Number of Hospitals Worked With & 1.63 & 1.48 & 1.29\\
\hspace{1em}Number of Systems Worked With & 1.38 & 1.30 & 1.16\\
\bottomrule
\end{tabular}
\label{tab:splitstats}
\end{table}