### 2022-08-04 Table Code/overall stats.tex

\begin{table}[ht]

\caption{Summary Statistics}
\centering
\begin{tabular}[t]{lccccc}
\toprule
Variable & N & Mean & Std. Dev. & Min & Max\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Outcomes}}\\
\hspace{1em}Ever Retire & 179,883 & 0.03 & 0.18 & 0 & 1\\
\hspace{1em}Fraction Patients in Office & 179,883 & 0.10 & 0.24 & 0 & 1\\
\hspace{1em}Work in an Office & 179,883 & 0.30 & 0.46 & 0 & 1\\
\hspace{1em}Change Zip Codes & 161,558 & 0.12 & 0.33 & 0 & 1\\
\hspace{1em}Number of Patients & 179,883 & 330 & 274 & 0 & 7,613\\
\hspace{1em}Claim Count & 179,883 & 1,382 & 1,789 & 0 & 64,508\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Treatment}}\\
\hspace{1em}Exposure to an EHR & 179,883 & 0.75 & 0.43 & 0 & 1\\
\hspace{1em}Fraction of Hospitals with EHR & 179,883 & 0.63 & 0.44 & 0 & 1\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Characteristics}}\\
\hspace{1em}Age & 179,883 & 44.9 & 8.4 & 25 & 86\\
\hspace{1em}Female & 179,883 & 0.34 & 0.47 & 0 & 1\\
\hspace{1em}Number of Hospitals Worked With & 179,883 & 1.49 & 0.82 & 1 & 12\\
\hspace{1em}Number of Systems Worked With & 179,883 & 1.31 & 0.60 & 1 & 8\\
\bottomrule
\end{tabular}
\label{tab:sumstats}
\end{table}