### 2022-08-23 Table Code/split stats.tex

\begin{table}[ht]

\caption{Means by Age Sample}
\centering
\begin{tabular}[t]{lcc}
\toprule
Variable & Age $>$ 60 & Age $<=$ 60\\
\midrule
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Outcomes}}\\
\hspace{1em}Ever Retire & 0.04 & 0.03\\
\hspace{1em}Fraction Patients in Office & 0.15 & 0.09\\
\hspace{1em}Work in an Office & 0.48 & 0.28\\
\hspace{1em}Change Zip Codes & 0.10 & 0.12\\
\hspace{1em}Number of Patients & 342 & 329\\
\hspace{1em}Claim Count & 1,767 & 1,332\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Treatment}}\\
\hspace{1em}Exposure to an EHR & 0.76 & 0.75\\
\hspace{1em}Fraction of Hospitals with EHR & 0.59 & 0.64\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Characteristics}}\\
\hspace{1em}Age & 60.31 & 42.75\\
\hspace{1em}Female & 0.18 & 0.36\\
\hspace{1em}Number of Hospitals Worked With & 1.63 & 1.48\\
\hspace{1em}Number of Systems Worked With & 1.38 & 1.30\\
\bottomrule
\end{tabular}
\end{table}
\label{tab:splitstats}