### 2022-09-10 Table Code/overall stats.tex

\begin{table}[ht]

\caption{Summary Statistics}
\centering
\begin{tabular}[t]{lccccc}
\toprule
Variable & N & Mean & Std. Dev. & Min & Max\\
\midrule
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Outcomes}}\\
\hspace{1em}Ever Retire & 162018 & 0.03 & 0.17 & 0 & 1.00\\
\hspace{1em}Fraction Patients in Office & 162018 & 0.08 & 0.21 & 0 & 1.00\\
\hspace{1em}Work in an Office & 162018 & 0.27 & 0.44 & 0 & 1.00\\
\hspace{1em}Change Zip Codes & 144787 & 0.12 & 0.32 & 0 & 1.00\\
\hspace{1em}Number of Patients & 162018 & 330.92 & 273.96 & 0 & 7613.00\\
\hspace{1em}Claims per Patient & 146443 & 4.09 & 4.09 & 0 & 480.68\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Treatment}}\\
\hspace{1em}Exposure to an EHR & 162018 & 0.76 & 0.43 & 0 & 1.00\\
\hspace{1em}Fraction of Hospitals with EHR & 162018 & 0.64 & 0.44 & 0 & 1.00\\
\addlinespace[0.3em]
\multicolumn{6}{l}{\textbf{Characteristics}}\\
\hspace{1em}Age & 162018 & 44.55 & 8.31 & 25 & 86.00\\
\hspace{1em}Female & 162018 & 0.34 & 0.47 & 0 & 1.00\\
\hspace{1em}Number of Hospitals Worked With & 162018 & 1.49 & 0.82 & 1 & 12.00\\
\hspace{1em}Number of Systems Worked With & 162018 & 1.31 & 0.61 & 1 & 8.00\\
\bottomrule
\end{tabular}
\end{table}
\label{tab:sumstats}