### 2023-04-20 Table Code/split stats.tex

\begin{table}[h]
\small
\caption{Means by Age, Exit Sample}
\centering
\begin{tabular}[t]{lcc}
\toprule
Variable & Age $<=$ 60 & Age $>$ 60\\
\midrule
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Outcomes}}\\
\hspace{1em}Ever Retire & 0.03 & 0.04\\
\hspace{1em}Work in an Office & 0.26 & 0.45\\
\hspace{1em}Fraction Patients in Office & 0.08 & 0.13\\
\hspace{1em}Number of Patients & 326.99 & 338.25\\
\hspace{1em}Claims Per Patient & 3.96 & 5.35\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Treatment}}\\
\hspace{1em}Year of EHR Exposure & 2010.82 & 2010.53\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Characteristics}}\\
\hspace{1em}Age & 42.60 & 60.24\\
\hspace{1em}Female & 0.36 & 0.18\\
\hspace{1em}Number of Hospitals Worked With & 1.46 & 1.62\\
\hspace{1em}Number of Systems Worked With & 1.29 & 1.38\\
\bottomrule
\end{tabular}
\end{table}