### 2024-03-20 Paper/coverletter.tex

\documentclass{article}
% Define WSU official colors
\usepackage[dvinames]{xcolor}
\definecolor{EmoryBlue}{RGB}{0,43,127} % Emory PMS 280
\definecolor{EmoryBlack}{RGB}{0,0,0}

% Measurements
\usepackage[top=2in,left=1in,bottom=0.5in,right=0.625in]{geometry}
\usepackage{graphicx}
\usepackage[colorlinks=false,
            pdfborder={0 0 0},
            ]{hyperref}
\usepackage[absolute]{textpos}
\usepackage{ifthen}
\usepackage{soul}

% --- A nice serif font (palatino), but not the prescribed nonfree ITC stone
\linespread{1.05}

% Remove paragraph indentation
\parindent0pt
\setlength{\parskip}{0.8\baselineskip}
\raggedright
\pagestyle{empty}
% Ensure consistency in the footer
\urlstyle{sf}

\providecommand\Emoryfrom{\textbf{Emory University}}
\providecommand\Emoryfrombuilding{Rich Memorial Building}
\providecommand\Emoryfromstreet{1602 Fishburne Drive, Room 306}
\providecommand\Emoryfromcity{Atlanta, GA 30322}
\providecommand\Emoryfromtel{404.727.6364 (tel)}
\providecommand\Emoryfromfax{404.727.4639 (fax)}
\providecommand\Emoryfromemail{\url{ian.mccarthy@emory.edu}}
\providecommand\Emorytoname{}
\providecommand\Emorytoaddress{}
\providecommand\Emoryre{}
\providecommand\Emorydate{\today}
\providecommand\Emoryopening{Dear Editors,}
\providecommand\Emoryfromname{Ian McCarthy}
\providecommand\Emoryfromtitle{Associate Professor}
\providecommand\Emoryclosing{Sincerely}

% Update this and the next line to the correct path
\providecommand\Emorysignaturefile{McCarthySignature.jpg}
\providecommand\Emorylogofile{EmoryLetterhead}
\providecommand\Emoryenclosure{}

\usepackage{fancyhdr}
\pagestyle{fancy}

\renewcommand{\footrulewidth}{0pt}
\fancyfoot{}
\fancyfoot[L]{%
    {\footnotesize\color{black}\sffamily
    \hspace{-.15in}
    \begin{tabular}{ll}
      \Emoryfrom & \Emoryfromtel \\[-0.1\baselineskip]
      \Emoryfrombuilding & \Emoryfromfax \\[-0.1\baselineskip]
      \Emoryfromstreet & \\[-0.1\baselineskip]
      \Emoryfromcity   &
    \end{tabular}
   }\color{black}}

\fancyhead{}
\fancyhead[L]{%
    \begin{textblock*}{2in}[0.3066,0.39](1.5in,1in)
        \includegraphics[width=2in]{EmoryLetterhead.pdf}
    \end{textblock*}
    \begin{textblock*}{4in}[0,0](3in,.7in)
      \begin{tabular}{|p{4in}}
        \vspace{.1cm}
        Ian M. McCarthy \\
        Associate Professor, Department of Economics \\
        Research Associate, National Bureau of Economic Research \\
        \vspace{0cm}
      \end{tabular}
    \end{textblock*}
    }
\renewcommand{\headrulewidth}{0pt}

\AtBeginDocument{
    % Text lines should be less than 6in long
    \newgeometry{top=1.8in,left=1in,bottom=1.2in,right=1in}

    \Emoryre
    \bigskip

    \Emorydate
    \bigskip

    \Emorytoname\ifthenelse{\equal{\Emorytoname}{}}{}{\\}
    \Emorytoaddress
    \bigskip

    \Emoryopening\par
    }

\AtEndDocument{
    \par\vspace{2ex}
    \Emoryclosing,

    \ifthenelse{\equal{\Emorysignaturefile}{}}{\bigskip\bigskip}{
      \includegraphics[width=2in]{\Emorysignaturefile}\\[-0.2\baselineskip]
    }

    \Emoryfromname \\
%    \Emoryfromtitle\ifthenelse{\equal{\Emoryfromtitle}{}}{}{\\}
    \Emoryenclosure
    }

\begin{document}
Enclosed please find our manuscript, ``Multimarket Contact in Health Insurance: Evidence from Medicare Advantage,'' to be considered for publication in the \textit{Journal of Industrial Economics}. This paper contributes to the literature on cross-market consolidations, which is of increasing interest to regulatory authorities and policy makers. We consider multimarket contact (MMC) as one potential mechanism by which cross-market consolidations can influence firm behaviors even in markets not directly affected by the merger.

We argue that Medicare Advantage is an excellent market to study these issues, not only because of the size and increasing reliance on Medicare Advantage to provide Medicare benefits, but also because of the transparency of potential deviations from collusive behavior and because of the availability of standardized quality measures made publicly available by CMS. Since our analysis is based on publicly available data, we are happy to provide all data and underlying code files should our paper be accepted. Thank you for your consideration, and we look forward to receiving your comments and suggestions.
\end{document}